Distributional Trees and Forests

### 2017-10-29 pkg/disttree/inst/draft/draft.Rnw
added text and edited plots
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1668 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 382285f56f532d0e6fa71d57578f3e26a9ac055e

\Abstract{
In the field of probabilistic forecasting classical linear models such as ensemble model output statistics (EMOS) are still common and widely used. This is due to their simplicity and their interpretability. However, the restriction to modelling only linear effects requires further extensions which allow for more flexibility. Generalized additive models for location, scale, and shape (GAMLSS, \cite{stasinopoulos2005}) can also capture non-linear effects as each distribution parameter can be modeled separately by its own GAM. A different approach that would make it possible to also model non-additive effects and interactions is to apply tree structured models. However, up to now, this class of models has not yet been integrated in probabilistic forecasting as no general distributional framework is available for these methods. Therefore, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. Applying these novel approaches on precipitation data in a mountainous region with a large number of numerical weather prediction quantities illustrates their strengths as their performance is compared to GAMLSS models that are specified based either on subject-matter knowledge/experience from the meteorology or based on a computationally more demanding boosting approach.

@@

\SweaveOpts{eval = TRUE}
\section{Introduction}
A typical application in probabilistic forecasting in meteorology is the analysis of precipitation data. The main two main interests lie in estimating the probability of rain as well as predicting the expected amount of precipitation. In this research area various methods have been developed and tested in the last decades. Basic linear models such as ensemble model output statistics (EMOS) were introduced in ... and are still common and widely used. Despite their simplicity and interpretability, the limitation to modelling only linear effects has risen the interest in applying more flexible models. This feature is provided by generalized additive models (GAM) which can also model non-linear effects. To obtain probabilistic predictions, generalized additive models for location, scale, and shape (GAMLSS) offer a framework in which each distribution parameter can be modeled separately by its own GAM. In Figure \ref{Devel_parmod} the development of parametric models starting from simple linear models (LM) and ranging to GAMLSS is illustrated.\\
These flexible parametric methods provide well fitting models when applied on precipitation data in a rather smooth/homogeneous setting, e.g. areas where no abrupt changes in meteorological/climatological events appear. However, for example in complex terrain/mountainous regions various factors can change rapidly and their effects are difficult to be modeled/represented by a smooth parametric model. In this situation a different approach which can deal with these abrupt changes would be to apply tree structured models. Moreover, these models are particularly advantageous in case of many covariates with unknown effects or interactions as the variable selection is done automatically within the model fitting process. Therefore, no additional effort or knowledge for the specification of the model in advance is required. However, even though their main advantage of being able to model non-additive effects and interactions without any necessary prespecification can play to their strength, regression trees are not an adequate choice if (additional) smooth effects appear as well. In this situation a random forest provides a good compromise as it combines a set of trees such that the steps imposed by the trees are smoothened in case of smooth effects while abrupt changes are still captured by the tree structure of the model.\\
For these reasons, regression trees and particularly random forests provide valuable features which can contribute to improving models for probabilistic forecasting. But, up to now, no general distributional framework is available for these methods. Therefore, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests.
(The developments in the field of tree-structured models together with the novel approaches of distributional trees and forests are illustrated in Figure \ref{Devel_treeforest}.)\\

@@

\newpage
\section{Methodology}
\label{Methodology}
\subsection{Distributional Fit}
Fitting a distributional model to a set of observations without considering any covariates is a well known procedure which can for example be done by applying the maximum likelihood method. This is also the chosen approach in the methodology described in this section.\\
The goal is to fit a distributional model $D(Y, \theta)$ to a response variable $Y \in \mathcal{Y}$ for a \textit{k}-dimensional vector of distribution parameters $\theta \in \Theta$ in case of a \textit{k}-parametric distribution. The distribution family has to be specified in advance such that a log-likelihood function $\ell(\theta; Y)$ is provided. In our application of precipitation forecasting the response variable, which is the power transformed total precipitation amount within 24 hours, is expected to be normally distribution with mean/location parameter $\mu$ and standard deviation/scale parameter $\sigma$ and left censored with censoring point 0. Therefore, the corresponding log-likelihood function $\ell_{tp}$ with parameter vector $\theta = c(\mu, \sigma)$ is
$$
\ell_{tp}(\mu, \sigma; Y) = 
\begin{cases}
    \log\left(\frac{1}{\sigma} \cdot \phi(\frac{Y - \mu}{\sigma}) \right), & \text{if } Y > 0\\
    \vspace{0.1cm}
    \log\left(\Phi(\frac{-\mu}{\sigma})\right), & \text{if } Y = 0
\end{cases}
$$
where $\phi$ and $\Phi$ are the probability density function and the distribution function of the standard normal distribution $\mathcal{N}(0,1)$.\\
With this specification the task of fitting a distributional model turns into the task of determining the distribution parameter $\theta$. This is done by the following maximization.
$$
\hat{\theta} = \max_{\theta \in \Theta} \sum_{i=1}^n \ell(\theta; y_i)
$$
where $\{y_i\}_{i=1,\ldots,n}$ are the observations of the response variable $Y$.\\
In that way, a whole distribution is specified with all its features, including location, scale and shape. However, fitting one global model to the whole data set might be too much of a generalization and therefore not represent the data and its features in a reasonable way. Moreover, if covariates are available it is desirable to include them into the model as they might provide important information. In particular, they can be used to separate the data into more homogeneous subgroups. Doing so after fitting a global model to the whole data set and then fitting a local model to each subgroup should improve the model remarkably. This procedure of splitting a data set into subgroups can be done by applying a tree algorithm.

@@

\SweaveOpts{eval = TRUE}
\newpage
\section{Probabilistic Forecasting of Precipitation}
As explained in the introduction distributional trees and forests provide a new approach in probabilistic forecasting which combines advantages of parametric models that are already used in this field and tree structured models which provide features that could be particularly valuable in probabilistic forecasting but have not yet been applied in this kind of settings. To show their strengths in practice, these new methods are now applied to probabilistic precipitation forecasts in a mountainous region based on a large number of numerical weather prediction quantities. In particular, the data set consists of daily 24h precipitation sums for the month of July from 28 years, starting in 1985 until 2012, both included. The observations were measured at 95 (?) stations all over Tyrol (and regions close to the frontier). The considered data is provided by ..... 
Next to the variable of interest, the measured total precipitation within 24 hours of each day (measured from 6 UTC to 6 UTC of the following day), numerical forecasts from the ECMWF are used as covariates. These forecasts are ensemble predictions consisting of ... individual forecasts based on perturbed initial conditions.\\
In particular, 14 basic forecast variables are considered with up to 12 variations of these variables being included as covariates. Since each forecast variable consists of an ensemble of individual predictions the possible variations include the mean, the standard deviation, the minimum and the maximum over all these individual predictions, noted as ensemble mean, ensemble std dev, ensemble min and ensemble max. All of these values can be calculated on the whole day (6--30 UTC) or a 6h time window of the day.\\
Another possible option is to consider the mean, minimum or maximum over 24 hours for each individual prediction and then take the ensemble mean or ensemble standard deviation, noted as ensemble mean of mean/min/max or ensemble std dev of mean/min/max respectively.\\
Table \ref{covariates} shows all variables together with the number of variations and the provided variations listed in the last column.\\
The response variable total precipitation is power transformed by a factor of 1.6. This is done due to experience and knowledge about possible distributions of precipitation measurements. The resulting transformed variable is expected to be normally distributed and left censored at censoring point 0.

@@

Next to distributional forests other methods / the same methods as in the simulation section are applied to this data set and their results are compared based on RMSE, log-likelihood values and CRPS values. The chosen methods and their specifications are listed in Table \ref{model_specification}.
\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
\begin{minipage}{\textwidth}
\begin{tabular}{| l | l   l   l |}
\hline
 & \textbf{Type} & \textbf{Location ($\mu$)} & \textbf{Scale ($\sigma$)} \\
\hline
\textbf{Forest} & recursive    & all & all \\ 
                & partitioning &     &     \\ 
\hline
\textbf{GAM (prespecified)} & spline  & tp{\_}mean, & tppow{\_}sprd,\\
                            & in each & tp{\_}max,  & dswrf{\_}sprd{\_}mean,\\
 & & tp{\_}mean1218 $\cdot$ & tp{\_}sprd1218 $\cdot$\\ 
 & & \quad cape{\_}mean1218, & \quad  cape{\_}mean1218,\\
 & & dswrf{\_}mean{\_}mean, & tcolc{\_}sprd{\_}mean,\\
 & & tcolc{\_}mean{\_}mean, & tdiff500850{\_}mean\\
 & & pwat{\_}mean{\_}mean, & \\
 & & tdiff500850{\_}mean, & \\
 & & msl{\_}diff & \\
\hline
\textbf{GAM (boosted)} & spline  & all & all \\
                       & in each &     &     \\
\hline
\textbf{EMOS} & linear & tp{\_}mean & tp{\_}spread \\
\hline
\end{tabular}
\caption[Table caption text]{Table of applied methods together with their type (way of including the covariates) and the included covariates for each distribution parameter}
\label{model_specification}
\end{minipage} 
%}
\end{table}

For the prespecified GAM model the covariates were selected manually based on meteorological knowledge and experience. Additionally, interactions are specified in advance, once between tp{\_}mean1218 and cape{\_}mean1218 for the location parameter and once between tp{\_}sprd1218 and cape{\_}mean1218 for the scale parameter. By putting extra effort in the specification of the formulas the prespecified GAM model is provided with additional information and is therefore expected to profit from this advantage.\\

\subsubsection{Means of Comparison}
In order to compare the performance of the chosen methods three different values are calculated for the resulting models: the root of the mean squared error (RMSE), the log-likelihood value and the continuous ranked probability score (CRPS). All of them are evaluated out of sample, e.g. on a part of the data set that has not been used to build the model. Therefore, overfitting would lead to worse results.\\
\\
The RMSE represents how close the estimated expected values are to the true observations. For a response variable $Y$ and a set of observations $\textbf{y} = \{y_i\}_{i = 1,\ldots,n}$ with the corresponding estimated parameters $\hat{\boldsymbol{\theta}} = \{\hat{\theta}_i\}_{i = 1,\ldots,n}$ the RMSE is defined as follows.

@@

In most cases this value does not measure the distributional fit as only the expected value is considered. However, in the case of a censored normal distribution the expected value of the observed variable is a transformation of the expected value of the latent variable which includes the location and the scale parameter.\\
\\
A typical way to evaluate the distributional fit is provided by the log-likelihood value. For each observation $y$ and the corresponding predicted parameter $\hat{\theta}$ the value of the log-likelihood function $\ell(\hat{theta}; y)$ is calculated and then averaged over all observations in the testing data set.\\
\\
The CRPS is another value that measures the accurarcy of a probabilistic model. For one observation $y$ with the predicted distribution function $F$ the CRPS value is calculated by
$$
\text{CRPS}(y, F) = \int_{-\infty}^\infty (F(z) - \mathbf{1}(y \le z))^2 dz
$$
where \(\mathbf{1}\) is the indicator function. Again, the value is averaged over all observations in the testing data set. As the CRPS can be seen as a generalization of the mean absolute error, a low value indicates a good fit. One big advantage in terms of interpretability is that the resulsting value is on the scale of the observation. \\
\\
As we are interested in comparing the performance of the distributional forest to those of other models, we choose one model as a reference and calculated the skills score with respect to the CRPS. In that way we obtain a measure of how much a method improves the model fit in terms of CRPS values compared to the reference method.
$$
\text{SkillsScore(method)} = 1 - \frac{\text{CRPS}_{\text{method}}}{\text{CRPS}_{\text{reference}}}
$$

\SweaveOpts{eval = TRUE}
\newpage
\subsection{Station Axams}
First of all we focus on one observation station, Axams in Tyrol. This station is selected due to its geographical closeness to Innsbruck, the capital of Tyrol. As for all other stations, daily observations of the month of July of 28 years are available, starting in 1985 and ending in 2012, both years included.\\ 
%\subsubsection{Predictions}
In a first step one model is learned on the first 25 years (1985--2009, both included) and then predictions for the 26th, 27th and 28th year (2010, 2011 and 2012) are made. To illustrate the results, one day of July is chosen and the estimated distributions for this day of all three prediction years are plotted together with the corresponding observations in Figure \ref{axams_onestation}.\\
The estimated distribution parameters for 24th of July of 2010 are comparably high with location $3.146$ and scale $1.902$ which leads to a well fitting prediction of $3.185$ close to the observed value of $3.001$. Also for 2011 and 2012 the estimated expected values of 1.085 and 0.483 provide reasonable predictions as the observed values are 1.121 and 0 respectively.\\
\\
<<echo=FALSE, results=hide>>=
#### Axams prediction 25 - 3
if(file.exists("rain_Axams_pred.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/rain_Axams_pred.rda")
  load("rain_Axams_pred.rda")
} else {
  #source("~/svn/partykit/pkg/disttree/inst/draft/rain_axams_pred.R")
  source("rain_axams_pred.R")
  library("gamlss.cens")
  gen.cens("NO", type = "left")
  res <- rain_axams_pred(seedconst = 7, ntree = 100,
                         forest_mtry = 27,   # if frac == TRUE: nvar/3 = 30
                         gamboost_cvr = TRUE,
                         frac = FALSE)

@@

#legendheight
lh1 <- crch::dcnorm(0.01, mean = df_mu[1], sd = df_sigma[1], left = 0)
lh2 <- crch::dcnorm(0.01, mean = df_mu[2], sd = df_sigma[2], left = 0)
lh3 <- crch::dcnorm(0.01, mean = df_mu[3], sd = df_sigma[3], left = 0)

\begin{figure}[t!]
<<rain_Axams_pred, fig=TRUE, echo=FALSE>>=
plot(x = x, y = y1, type = "l", col = "red", 
     main = paste0("July ", pday, dayending), ylab = "density", 
     xlab = "total precipitation (power transformed)",
     ylim = c(0,max(y1, y2, y3, pm1, pm2, pm3) + 0.01),
     xlim = c(-1.5,8))

@@

    # add labels
    text(x = -0.8, y = lh1, labels = "2010", col = "red", cex = 0.8)
    text(x = -0.8, y = lh2, labels = "2011", col = "blue", cex = 0.8)
    text(x = -0.8, y = lh3, labels = "2012", col = "darkgreen", cex = 0.8)
@
\caption{\label{axams_oneday}Predictions at station Axams for one day in July of 2010, 2011 and 2012}
\end{figure}

Analyzing the distforest model in this setting it can be seen that the covariates listed in the following plot show the highest values in variable importance.\\
As expected, the numerical forecasts variable for total precipitation is the most frequently selected. In particular, three variations of this variable lead the top ten of this list.\\  
<<variabble_importance, echo=FALSE, eval=TRUE>>=
if(file.exists("vimp_crps.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/vimp_crps.rda")
  load("vimp_crps.rda")
} else {

@@

\begin{figure}[t!]
%\begin{subfigure}{0.5\textwidth}
<<rain_axams_varim_margeffects1_mu, fig=TRUE, echo=FALSE>>=
par(mar = c(2,10,2,2))
barplot(sort(vimp_crps, decreasing = TRUE)[1:10], horiz = TRUE, las = 1, cex.names = 0.8, axes = FALSE)
@
%\end{subfigure}
%\hspace{-1cm}
%\begin{subfigure}{0.5\textwidth}
%<<rain_axams_varim_margeffects1_sigma, fig=TRUE, echo=FALSE>>=
%plot(x = sp1[,id_mi[1]], y = sp1[, "sigma"], type = "l", xlab = names_mi[1], ylab = %"sigma")
%@
%\end{subfigure}
\caption{\label{varimp} Variable importance: top 10 list for the forest model}
\end{figure}

To get another impression of the quality of the predictions made by the tested methods two PIT-histograms are plotted for each of them, one for the in-sample predictions and one for the out-of-sample predictions.\\

<<pit_qqr, echo=FALSE, eval=TRUE>>=

@@

  ## disttree
  # in sample
  pdt <- predict(res$dt, type = "parameter")
  dt_mu_l <- pdt$mu 
  dt_sigma_l <- pdt$sigma 
  pit_dt_l <- cbind(0, pnorm(res$learndata[,"robs"], mean = dt_mu_l, sd = dt_sigma_l))
  pit_dt_l[res$learndata[,"robs"]>0, 1] <- pit_dt_l[res$learndata[,"robs"]>0, 2]

  # out of sample
  pdt <- predict(res$dt, type = "parameter", newdata = res$testdata)
  dt_mu_t <- pdt$mu 
  dt_sigma_t <- pdt$sigma 
  pit_dt_t <- cbind(0, pnorm(res$testdata[,"robs"], mean = dt_mu_t, sd = dt_sigma_t))
  pit_dt_t[res$testdata[,"robs"]>0, 1] <- pit_dt_t[res$testdata[,"robs"]>0, 2]

  ## distforest
  # in sample
  pdf <- predict(res$df, type = "parameter")
  df_mu_l <- pdf$mu
  df_sigma_l <- pdf$sigma
  pit_df_l <- cbind(0, pnorm(res$learndata[,"robs"], mean = df_mu_l, sd = df_sigma_l))
  pit_df_l[res$learndata[,"robs"]>0, 1] <- pit_df_l[res$learndata[,"robs"]>0, 2]

>>>>

\Abstract{
In the field of probabilistic forecasting classical linear models such as ensemble model output statistics (EMOS, \cite{gneiting2005calibrated}) provide common and widely used tools. This is due to their simplicity and their interpretability. However, the restriction to modeling only linear effects requires further extensions which allow for more flexibility. Generalized additive models for location, scale, and shape (GAMLSS, \cite{stasinopoulos2005}) can also capture non-linear effects as each distribution parameter can be modeled separately by its own generalized additive model (GAM). A different approach that would make it possible to also model non-additive effects and interactions is to apply tree structured models. However, up to now, this class of models has not yet been integrated in probabilistic forecasting as no general distributional framework is available for these methods. Therefore, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. Applying these novel approaches on precipitation data in a mountainous region with a large number of numerical weather prediction quantities illustrates their strengths as their performance is compared to GAMLSS models that are specified based either on subject-matter knowledge/experience from the meteorology or based on a computationally more demanding boosting approach.

@@

\SweaveOpts{eval = TRUE}
\section{Introduction}
A typical application in probabilistic forecasting in meteorology is the analysis of precipitation data. The two main interests lie in estimating the probability of rain as well as predicting the expected amount of precipitation. In this research area various methods have been developed and tested in the last decades. Basic linear models such as ensemble model output statistics (EMOS) were introduced more than a decade ago (\cite{gneiting2005calibrated}) and are still common and widely used. Despite their simplicity and interpretability, the limitation to modeling only linear effects has risen the interest in applying more flexible models. This feature is provided by generalized additive models (GAM) which can also model non-linear effects. However, often only the expectation of the underlying distribution is estimated. To obtain probabilistic predictions, generalized additive models for location, scale, and shape (GAMLSS) offer a framework in which each distribution parameter can be modeled separately by its own GAM. In Figure \ref{Devel_parmod} the development of parametric models starting from simple linear models (LM) and ranging to GAMLSS is illustrated.\\
These flexible parametric methods provide well fitting models when applied on precipitation data in a rather smooth/homogeneous setting, e.g. areas where no abrupt changes in meteorological/climatological events appear. However, for example in complex terrain/mountainous regions various factors can change rapidly and their effects are difficult to be modeled/represented by a smooth parametric model. In this situation a different approach which can deal with these abrupt changes would be to apply tree structured models. Moreover, these models are particularly advantageous in case of many covariates with unknown effects or interactions as the variable selection is done automatically within the model fitting process. Therefore, no additional effort or knowledge for the specification of the model in advance is required. However, even though their main advantage of being able to model non-additive effects and interactions without any necessary prespecification can play to their strength, regression trees are not an adequate choice if (additional) smooth effects appear as well. In this situation a random forest provides a good compromise as it combines a set of trees such that the steps imposed by the trees are smoothened in case of smooth effects while abrupt changes are still captured by the tree structure of the model.\\
For these reasons, regression trees and particularly random forests provide valuable features which can contribute to improving models for probabilistic forecasting. But, up to now, no general distributional framework is available for these methods. Therefore, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. The developments in the field of tree-structured models together with the novel approaches of distributional trees and forests are illustrated in Figure \ref{Devel_treeforest}.\\

@@

\newpage
\section{Methodology}
\label{Methodology}
\subsection{Distributional Fit}
Fitting a distributional model to a set of observations without considering any covariates is a well known procedure which can for example be done by applying the maximum likelihood method. This is also the chosen approach in the methodology described in this section.\\
The goal is to fit a distributional model $D(Y, \theta)$ to a response variable $Y \in \mathcal{Y}$ for a \textit{k}-dimensional vector of distribution parameters $\theta \in \Theta$ in case of a \textit{k}-parametric distribution. The distribution family has to be specified in advance such that a log-likelihood function $\ell(\theta; Y)$ is provided. In our application of precipitation forecasting the response variable, which is the power transformed total precipitation amount within 24 hours, is expected to be normally distribution with mean/location parameter $\mu$ and standard deviation/scale parameter $\sigma$ and left censored with censoring point 0. Therefore, the corresponding log-likelihood function $\ell_{tp}$ with parameter vector $\theta = (\mu, \sigma)$ is
$$
\ell_{tp}(\mu, \sigma; Y) = 
\begin{cases}
    \log\left(\frac{1}{\sigma} \cdot \phi(\frac{Y - \mu}{\sigma}) \right), & \text{if } Y > 0\\
    \vspace{0.1cm}
    \log\left(\Phi(\frac{-\mu}{\sigma})\right), & \text{if } Y = 0
\end{cases}
$$
where $\phi$ and $\Phi$ are the probability density function and the distribution function of the standard normal distribution $\mathcal{N}(0,1)$.\\
With this specification the task of fitting a distributional model turns into the task of determining the distribution parameter $\theta$. This is done by the following maximization.
$$
\hat{\theta} = \max_{\theta \in \Theta} \sum_{i=1}^n \ell(\theta; y_i)
$$
where $\{y_i\}_{i=1,\ldots,n}$ are the observations of the response variable $Y$.\\
In that way, a whole distribution is specified with all its features, including location, scale and shape. However, fitting one global model to the whole data set might be too much of a generalization and therefore not represent the data and its features in a reasonable way. Moreover, if covariates are available it is desirable to include them into the model as they might provide important information. In our application on precipitation data a wide range of covariates are available from numerical esemble predictions considering precipitation, temperature, sea level pressure and other meteorological quantities which might be of interest/have an influence on the predictions. In particular, these covariates can be used to separate the data into more homogeneous subgroups. Doing so after fitting a global model to the whole data set and then fitting a local model to each subgroup should improve the model remarkably. This procedure of splitting a data set into subgroups can be done by applying a tree algorithm.

@@

\SweaveOpts{eval = TRUE}
\newpage
\newpage
\section{Probabilistic Forecasting of Precipitation}
As explained in the introduction distributional trees and forests provide a new approach in probabilistic forecasting which combines advantages of parametric models that are already used in this field and tree structured models which provide features that could be particularly valuable in probabilistic forecasting but have not yet been applied in this kind of settings. To show their strengths in practice, these new methods are now applied to probabilistic precipitation forecasts in a mountainous region based on a large number of numerical weather prediction quantities. In particular, the data set consists of daily 24h precipitation sums for the month of July from 28 years, starting in 1985 until 2012, both included. The observations were measured at 95 (?) stations all over Tyrol (and regions close to the frontier). The considered data is provided by ..... 
Next to the variable of interest, the measured total precipitation within 24 hours of each day (measured from 6 UTC to 6 UTC of the following day), numerical forecasts from the ECMWF are used as covariates. These forecasts are ensemble predictions consisting of ... individual forecasts based on perturbed initial conditions.\\
In particular, 14 basic forecast variables are considered with up to 12 variations of each of these variables being included as covariates. Since each forecast variable consists of an ensemble of individual predictions the possible variations include the mean, the standard deviation, the minimum and the maximum over all these individual predictions, noted as ensemble mean, ensemble std dev, ensemble minimum and ensemble maximum. All of these values can be calculated on the whole day (6--30 UTC) or a 6h time window of the day.\\
Another possible option is to consider the mean, minimum or maximum over 24 hours for each individual prediction and then take the ensemble mean or ensemble standard deviation, noted as ensemble mean of mean/minimal/maximal values or ensemble std dev of mean/minimal/maximal values respectively.\\
Table \ref{covariates} shows all variables together with the number of variations and the provided variations listed in the last column. Altogether 82 covariates are available in this setting.\\
The response variable, total precipitation (tp), is power transformed by a factor of 1.6. This is done due to experience and knowledge about possible distributions of precipitation measurements. The resulting transformed variable is expected to be normally distributed and left censored at censoring point 0.

@@

Next to distributional forests other methods / the same methods as in the simulation section are applied to this data set and their results are compared based on RMSE, log-likelihood values and CRPS values. The chosen methods and their specifications are listed in Table \ref{model_specification}.
\begin{table}[t!]
%\resizebox{1.4\textwidth}{!}{
%\begin{minipage}{\textwidth}
\begin{center}
\begin{tabular}{| l | l   l   l |}
\hline
 & \textbf{Type} & \textbf{Location ($\boldsymbol{\mu}$)} & \textbf{Scale ($\boldsymbol{\log(\sigma)}$)} \\
\hline
\textbf{Forest} & recursive    & all & all \\ 
                & partitioning &     &     \\ 
\hline
\textbf{GAM (prespecified)} & spline  & tp{\_}mean, & tppow{\_}sprd,\\
                            & in each & tp{\_}max,  & dswrf{\_}sprd{\_}mean,\\
 & & tp{\_}mean1218 $\cdot$ & tp{\_}sprd1218 $\cdot$\\ 
 & & \quad cape{\_}mean1218, & \quad  cape{\_}mean1218,\\
 & & dswrf{\_}mean{\_}mean, & tcolc{\_}sprd{\_}mean,\\
 & & tcolc{\_}mean{\_}mean, & tdiff500850{\_}mean\\
 & & pwat{\_}mean{\_}mean, & \\
 & & tdiff500850{\_}mean, & \\
 & & msl{\_}diff & \\
\hline
\textbf{GAM (boosted)} & spline  & all & all \\
                       & in each &     &     \\
\hline
\textbf{EMOS} & linear & tp{\_}mean & tp{\_}spread \\
\hline
\end{tabular}
\caption[Table caption text]{Table of applied methods together with their type (way of including the covariates) and the included covariates for each distribution parameter}
\label{model_specification}
%}
\end{center}
\end{table}

For the prespecified GAM model the covariates were selected manually based on meteorological knowledge and experience. Additionally, interactions are specified in advance, once between tp{\_}mean1218 and cape{\_}mean1218 for the location parameter and once between tp{\_}sprd1218 and cape{\_}mean1218 for the scale parameter. By putting extra effort in the specification of the formulas the prespecified GAM model is provided with additional information and is therefore expected to profit from this advantage.\\
In the EMOS model the logarithmic function is chosen as the link function for the scale parameter. This choice has been made after testing and comparing three link functions (identity, quadratic and logarithmic function) by selecting the one leading to the best performance of the EMOS model in this application.

\subsubsection{Means of Comparison}
In order to compare the performance of the chosen methods three different values are calculated for the resulting models: the root of the mean squared error (RMSE), the log-likelihood value and the continuous ranked probability score (CRPS). All of them are evaluated out of sample, e.g. on a part of the data set that has not been used to build the model.\\
%Therefore, overfitting would lead to worse results.\\
\\
The RMSE represents how close the estimated expected values are to the true observations. For a response variable $Y$ and a set of observations $\textbf{y} = \{y_i\}_{i = 1,\ldots,n}$ with the corresponding estimated parameters $\hat{\boldsymbol{\theta}} = \{\hat{\theta}_i\}_{i = 1,\ldots,n}$ the RMSE is defined as follows.

@@

In most cases this value does not measure the distributional fit as only the expected value is considered. However, in the case of a censored normal distribution the expected value of the observed variable is a transformation of the expected value of the latent variable which includes the location and the scale parameter.\\
\\
A typical way to evaluate the distributional fit is provided by the log-likelihood value. For each observation $y$ and the corresponding predicted parameter $\hat{\theta}$ the value of the log-likelihood function $\ell(\hat{\theta}; y)$ is calculated and then averaged over all observations in the testing data set. The higher the value the better the predictions of the model are.\\
\\
The CRPS is another value that measures the accuracy of a probabilistic model. For one observation $y$ with the predicted distribution function $F$ the CRPS value is calculated by
$$
\text{CRPS}(y, F) = \int_{-\infty}^\infty (F(z) - \mathbf{1}(y \le z))^2 dz
$$
where \(\mathbf{1}\) is the indicator function. Again, the value is averaged over all observations in the testing data set. As the CRPS can be seen as a generalization of the mean absolute error, a low value indicates a good fit. One big advantage in terms of interpretability is that the resulting value is on the scale of the observation. \\
\\
As we are interested in comparing the performance of the distributional forest to those of other models, we choose one model as a reference and calculated the skills score with respect to the CRPS. In that way we obtain a measure of how much a method improves the model fit in terms of CRPS values compared to the reference method.
$$
\text{SkillsScore(method)} = 1 - \frac{\text{CRPS}_{\text{method}}}{\text{CRPS}_{\text{reference}}}
$$

\SweaveOpts{eval = TRUE}
\subsection{Station Axams}
First of all we focus on one observation station, Axams in Tyrol. This station is selected due to its geographical closeness to Innsbruck, the capital of Tyrol. As for all other stations, daily observations of the month of July of 28 years are available, starting in 1985 and ending in 2012, both years included.\\ 
%\subsubsection{Predictions}
In a first step one model is learned on the first 25 years (1985--2009, both included) and then predictions for the 26th, 27th and 28th year (2010, 2011 and 2012) are made. To illustrate the results, one day of July is chosen and the density functions of the estimated distributions for this day of all three prediction years are plotted together with the corresponding observations in Figure \ref{axams_onestation}.\\
The estimated distribution parameters for the 24th of July of 2010 are comparably high with location $3.146$ and scale $1.902$ which leads to a well fitting predicted expectation of $3.185$ close to the observed value of $3.001$. Also for 2011 and 2012 the estimated expected values of 1.085 and 0.483 provide reasonable predictions as the observed values are 1.121 and 0 respectively.\\

<<echo=FALSE, results=hide>>=
#### Axams prediction 25 - 3
if(file.exists("rain_Axams_pred.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/rain_Axams_pred.rda")
  load("rain_Axams_pred.rda")
} else {
  #source("~/svn/partykit/pkg/disttree/inst/draft/rain_axams_pred.R")
  source("rain_axams_pred.R")
  library("gamlss.cens")
  gen.cens("NO", type = "left")
  res <- rain_axams_pred(seedconst = 7, ntree = 100,
                         forest_mtry = 27,   # if frac == TRUE: nvar/3 = 30
                         gamboost_cvr = TRUE,
                         frac = FALSE)

@@

#legendheight
lh1 <- crch::dcnorm(0.01, mean = df_mu[1], sd = df_sigma[1], left = 0)
lh2 <- crch::dcnorm(0.01, mean = df_mu[2], sd = df_sigma[2], left = 0)
lh3 <- crch::dcnorm(0.01, mean = df_mu[3], sd = df_sigma[3], left = 0)
@

\begin{figure}[t!]
\begin{center}
<<rain_Axams_pred, fig=TRUE, echo=FALSE>>=
plot(x = x, y = y1, type = "l", col = "red", 
     main = paste0("July ", pday, dayending), ylab = "density", 
     xlab = "total precipitation (power transformed)",
     ylim = c(0,max(y1, y2, y3, pm1, pm2, pm3) + 0.01),
     xlim = c(-1.5,8))

@@

    # add labels
    text(x = -0.8, y = lh1, labels = "2010", col = "red", cex = 0.8)
    text(x = -0.8, y = lh2, labels = "2011", col = "blue", cex = 0.8)
    text(x = -0.8, y = lh3, labels = "2012", col = "darkgreen", cex = 0.8)
@
\caption{\label{axams_onestation}Predictions at station Axams for one day in July of 2010, 2011 and 2012}
\end{center}
\end{figure}

As seen in Table \ref{model_specification} the forest model includes all 82 covariates. In a single tree model it can easily be observed which covariates are used as split variables and hence have an impact on the response variable. In a forest model though, for each of the trees a subset of these covariates is chosen randomly as possible split variables which is one of the reasons leading to different splitting variables being chosen in each tree. To get an overview of which covariates have the most influential effect on the response variable a standard measurement of variable importance for forest models has been chosen. The 10 leading covariates of the above described forest model for the station Axams learned on the first 25 year are listed in Figure \ref{varimp}. As expected, the numerical forecast for total precipitation is the most frequently selected. In particular, five variations of this variable are among the top ten of the list of covariates ordered by variable importance.  
<<variabble_importance, echo=FALSE, eval=TRUE>>=
if(file.exists("vimp_crps.rda")){
  #load("~/svn/partykit/pkg/disttree/inst/draft/vimp_crps.rda")
  load("vimp_crps.rda")
} else {

@@

\begin{figure}[t!]
\begin{center}
\begin{subfigure}{0.7\textwidth}
<<rain_axams_varim, fig=TRUE, echo=FALSE>>=
par(mar = c(2,10,2,2))
barplot(sort(vimp_crps, decreasing = TRUE)[1:10], horiz = TRUE, las = 1, cex.names = 1.2, axes = FALSE, names.arg = gsub("pow", "", names(sort(vimp_crps, decreasing = TRUE)[1:10])))
@
\end{subfigure}
%\hspace{-1cm}
%\begin{subfigure}{0.5\textwidth}
%<<rain_axams_varim_margeffects1_sigma, fig=TRUE, echo=FALSE>>=
%plot(x = sp1[,id_mi[1]], y = sp1[, "sigma"], type = "l", xlab = names_mi[1], ylab = %"sigma")
%@
%\end{subfigure}
\caption{\label{varimp} Variable importance: top 10 list for the forest model}
\end{center}
\end{figure}

To get another impression of the quality of the predictions made by the tested methods a probability integral transform (PIT) histograms is plotted in Figure \ref{pit_oosample} for out-of-sample predictions of each of the models. A perfectly fitting model would lead to a histogram with all bars being of height 1. Overall, it can be stated that all four model provide well fitting models while the forest seems to be the one being closest to the ideal case.

<<pit_qqr, echo=FALSE, eval=TRUE, results=hide>>=

@@

  ## disttree
  if(FALSE){
    # in sample
    pdt <- predict(res$dt, type = "parameter")
    dt_mu_l <- pdt$mu 
    dt_sigma_l <- pdt$sigma 
    pit_dt_l <- cbind(0, pnorm(res$learndata[,"robs"], mean = dt_mu_l, sd = dt_sigma_l))
    pit_dt_l[res$learndata[,"robs"]>0, 1] <- pit_dt_l[res$learndata[,"robs"]>0, 2]
  }  
  # out of sample
  pdt <- predict(res$dt, type = "parameter", newdata = res$testdata)
  dt_mu_t <- pdt$mu 
  dt_sigma_t <- pdt$sigma 
  pit_dt_t <- cbind(0, pnorm(res$testdata[,"robs"], mean = dt_mu_t, sd = dt_sigma_t))
  pit_dt_t[res$testdata[,"robs"]>0, 1] <- pit_dt_t[res$testdata[,"robs"]>0, 2]
